--- 
title: "Estimating Acid Soil Effects on Cereal Crop Productivity in Ethiopia: A Digital Handbook "
author: "Michael Mann, James Warner, Chilot Tizale, Jordan Chamberlin"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: ""
---

# Index {-}  

The objective of this digital handbook is to provide insight into the methodology and outcomes of our paper "Estimating acid soil effects on cereal crop productivity in Ethiopia: A prototype for enhancing crop management". 

- **Chapter 1:** [Introduction](#intro)
- **Chapter 2:** [Data](#data)
- **Chapter 3:** [Methods](#methods)
- **Chapter 4:** [Findings](#findings)
- **Chapter 5:** [Conclusions](#conclusions)


***


**Tutorials**

We provide a small set of tutorials to help provide insight into the methods used in this paper.

- **Tutorial 1:** [GAMS Tutorial](#gams)


***


**Policy Maps**

To help guide agricultural interventions we provide a set of maps. These can be used to help better targeting of liming and fertilizer interventions. 

- **Maps 1:** [Local Estimates](#local)


```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
library(sf)
library(broom)
library(car)
library(raster)
library(readstata13)
library(dplyr)
library(plm)
library(splines2)
require(stats)
library(splines)
library(ggplot2)
library(reshape2)
library(stars)
library(pdp)
library(pander)
library(caret)
library(FinCal)
library(tidyr)
library(automap)
library(flextable)
library(officer)
library(magrittr)
library(rms)
library(tidyselect)
library(english)
library(readstata13)
library(tmap)    # for static and interactive maps
library(leaflet) # for interactive maps
library(mapview) # for interactive maps

# set up numbering for figures/graphs etc
Table_number = 1
Figure_number = 1
Formula_number = 1
Foot_number = 1


knitr::opts_chunk$set(
  echo = F,
  message = F,
  #collapse = TRUE,
  warning = F
  #comment = "#>", 
  #eval = !is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))
)

# standard boarder for flextable
std_border = fp_border(color="grey", width = 1)




winsorize_x = function(x, cut = 0.01){
  cut_point_top <- quantile(x, 1 - cut, na.rm = T)
  #cut_point_bottom <- quantile(x, cut, na.rm = T)
  i = which(x >= cut_point_top) 
  x[i] = cut_point_top
  #j = which(x <= cut_point_bottom) 
  #x[j] = cut_point_bottom
  return(x)
}

```


```{r cost analysis assumptions, include=FALSE}
# Here are some fertilizer prices by region for 2018.  I think we should use the Urea and NPS (the NPS has pretty much replaced DAP going forward).
# 
# Per 100 kgs (in Birr)
#                     Dap       Urea     NPS
# Tigray              1495      1327      1383
# Amhara              1353      1229      1413
# Oromia              1632      1272      1397
# SNNP                1228      1157      1220


price = 2000
price_dollars = round(price/27.52)
fertilizer_P_kg= 15 #00 # mean(1382,1413,1397,1220) /100   # 1500 set by james
tons_lime_ha = 3
total_lime_cost = price * tons_lime_ha

# discounting parameters
discount_rate = 0.15
pv_years = 5

# 1 million hectares of abandoned land to lime


```


```{r include=FALSE}
# load results from conclusion to insert into abstract
h_acidic_oph = readRDS('./Images/h_acidic_oph.rds')
ph_oph = readRDS('./Images/PH_dataframe.rds')

```
